---
title: "Workshop 1"
author: "Aprendizaje Estadístico 2"
format: pdf
editor: visual
---

Solve the next exercises using computational tools like R or Python, R Mardown or Google Colab. Make a pdf report where you show the whole procedure. Take into account that you have to be clear because it is something to be assesed.

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)
```

```{r}
library(tidyverse)
```

## Experiment Setup

From exercise 1 to 5 consider the sample space of tossing a fair coin 3 times:

$$S=\{HHH,HHT,HTH,THH,HTT,THT,TTH,TTT\}$$

with each outcome having probability $1/8$.

1.  Describe the next events:

$A:$ "first toss is $H$." $B:$ "second toss is $H$." $C:$ "all tosses are $H$." $D:$ "exactly one $H$."

2.  Check if $A$ and $B$ are independent. Are they mutually exclusive?

    ```{r}

    ```

3.  Check if $C$ and $D$ are independent. Are they mutually exclusive?

    ```{r}

    ```

4.  Prove: If two events $E, F$ are mutually exclusive and both have positive probability, then they cannot be independent.

    ```{r}

    ```

5.  Take event $G=$ “first toss is head,” $G^C=$ “first toss is tail.” Show $G$ and $G^C$ are mutually exclusive and compute whether they are independent.

    ```{r}

    ```

6.  Compute $P(Z \ge 1.96)$ for a standard normal $Z \sim \mathcal N(0,1)$ and illustrate it by shading the right tail.

Using pnorm(), compute $P(Z \ge 1.96)$.

```{r}

```

Make a plot of the standard normal density on $x \in [-4,4]$ with the right tail ($x \ge 1.96$) shaded.

```{r}

```

7.  For $Z \sim \mathcal N(0,1)$ compute:

<!-- -->

(a) $P(Z \le -1.5)$

    ```{r}

    ```

(b) $P(Z \ge 2.1)$

    ```{r}

    ```

(c) $P(|Z| \ge 2)$

    ```{r}

    ```

(d) The central 95% probability $P(-z \le Z \le z)$ and the value of $z$ such that this equals $0.95$.

    ```{r}

    ```

<!-- -->

8.  dd

    (a) Find the 90th percentile of $Z \sim \mathcal N(0,1)$, i.e., compute $z_{0.90}$ such that $P(Z \le z_{0.90})=0.90$.

        ```{r}

        ```

    (b) Find $z$ such that $P(|Z| \le z)=0.90$.

        ```{r}

        ```

    (c) Suppose $X \sim \mathcal N(\mu=100, \sigma=15)$. Compute the score threshold $c$ for the top 5%.

        ```{r}

        ```

9.  For $Z \sim \mathcal N(0,1)$:

<!-- -->

(a) Simulate $N=100{,}000$ draws and estimate $P(Z \ge 1.96)$ empirically.

    ```{r}
    x <- rnorm(100, 5, 2)
    mean(x)
    hist(x)
    ```

(b) Compare with the exact value.

    ```{r}

    ```

(c) Plot a histogram with an overlaid theoretical density curve.

    ```{r}

    ```

<!-- -->

10. Let $X \sim \mathcal N(\mu=70,\ \sigma=8)$. Compute:

<!-- -->

(a) $P(X \le 60)$

    ```{r}

    ```

(b) $P(65 \le X \le 85)$

    ```{r}

    ```

(c) The 2.5th percentile of $X$

    ```{r}

    ```

Do each both directly and by standardizing to $Z=\dfrac{X-\mu}{\sigma}$.

11. Generate $N=200$ observations from $\mathcal N(\mu = 5, \sigma^2=2^2)$, then:

    ```{r}

    ```

<!-- -->

(a) Plot histogram with density overlay.

    ```{r}

    ```

(b) Make a QQ‑plot to assess normality.

    ```{r}

    ```

<!-- -->

12. 

    (a) For $Z \sim \mathcal N(0,1)$, compute $z_\alpha$ such that $P(|Z| \le z_\alpha) = 0.99$.

        ```{r}

        ```

    (b) For $X \sim \mathcal N(\mu = 120, \sigma=10)$, find the symmetric 90% interval around the mean.

    (c) 

        ```{r}

        ```

    (d) Verify by simulation that about 90% of draws fall in your interval from (b).

        ```{r}

        ```
